---
title: "Red Wine EDA Report"
author: "Yousef Alotaibi"
date: '2022-06-14'
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../"
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Libraries
library(tidyverse)
library("janitor")
library(reshape2)
library(caret) # for multiple box plot
library("PerformanceAnalytics") # Correlation matrix
library(here)
library(scales)
library(DT)

# reading the dataset
wineRed <- read_csv(here('data/winequality-red-kaggle.csv'))

```

## Atrributes description 
+--------------------+---------------+----------------------------------------------+
|Variable            |Type           |    Description                               |
+====================+===============+==============================================+
| fixed_acidity      |    Double     | most acids involved with wine or fixed or    |
|                    |               | nonvolatile (do not evaporate readily)       |
+--------------------+---------------+----------------------------------------------+
|Voltaile_acidity    |    Double     | the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
| citric_acid        |    Double     | found in small quantities, citric acid can add 'freshness' and flavor to wines|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
|residual_sugar      |    Double     | the amount of sugar remaining after fermentation stops|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
|chlorides           |    Double     | the amount of salt in the wine               |
+--------------------+---------------+----------------------------------------------+
|free_sulfur_dioxide |    Double     | the free form of SO2 exists in equilibrium between molecular SO2| 
|                    |               | (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
|total_sulfur_dioxide|    Double     | amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm,| 
|                    |               | SO2 becomes evident in the nose and taste of wine|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
|density             |    Double     | the density of water is close to that of water depending on the percent alcohol and sugar content|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
|pH                  |    Double     | describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
|sulphates           |    Double     | a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant|
|                    |               |                                              |
+--------------------+---------------+----------------------------------------------+
|alcohol             |    Double     | the percent alcohol content of the wine      |
+--------------------+---------------+----------------------------------------------+
|quality             |    Double     | Score between (3-8)                          | 
+--------------------+---------------+----------------------------------------------+

## Interactive table 
```{r}
datatable(wineRed)

```

```{r, echo=FALSE, include=FALSE}

head(wineRed)

wineRed <- rename_with(wineRed, ~ tolower(gsub(" ", "_", .x, fixed = TRUE)))

get_dupes(wineRed)
# We can see that we have 230 duplicated observations.
# Let's remove them! 

wineRed <- wineRed[!duplicated(wineRed), ]
# New tibble with only non-duplicate observations.

is.null(wineRed)
# checking for null values 

summary(wineRed)

qualityCat <- cut(wineRed$quality, 
                  breaks=c(0,4,5,6,7,8), 
                  labels=c("undrinkable (2-4)",
                  "pretty bad (5)",
                  "fair, but has noticable flaws (6)",
                  "quaffable (7)",
                  "very good (8)"))
```



```{r}
ggplot(data = wineRed, aes(x = qualityCat)) +
  geom_bar(fill='light blue') +
geom_text(
     aes(label= scales::percent(after_stat(as.double(prop))), group=1),
     stat='count', vjust = -0.3, 
 ) 
 
```

#### Histogram shows us that wine quality between 7-8 is around 14% of our dataset, 40% is 6 and 46% between 2-5. 

```{r}
cor(wineRed[-1], wineRed$quality)

```
#### - quality has a large positive correlation with alcohol,
#### a moderate positive correlation with sulphates 
#### and a moderate negative correlation with volatile_acidity.
``` {r}
chart.Correlation(wineRed, histogram=TRUE, pch=19)
```

#### - It's fair to say that density and pH are normally
#### distributed, which means observations have a high frequency near the mean.
#### - fixed_acidity and citric_acid have a large positive correlation(0.67).
#### - fixed_acidity and density have a large positive corrrelation(0.67).
#### - fixed_acidity and pH have a large negative correlation(-0.68).
#### - citric_acid and pH have a large negative correlation(-0.54).

```{r, fig.width = 15, fig.height=15}
featurePlot(x = wineRed[, 1:11], 
            y = qualityCat, 
            plot = "box", 
            scales = list(y = list(relation="free"),
                          x = list(rot = 90)),  
            layout = c(3,4), 
            auto.key = list(columns = 2))
```

#### all of the variables affect in increasing or decreasing the wine quality, excluding chlorides
#### and residual_sugar as they have equal median for all qualities.

## Conclusion 

#### In conclusion I have found that wine quality is likely to be high due to the affect of the following      variables:   
####              - high levels of alcohol results in high wine quality.
####              - increasing sulphates and fixed_acidity will moderately increase wine quality.
####              - low levels of volatile_acidity results in high wine quality.
####              - Previously we mentioned that pH is negatively correlated with both fixed_acidity and citric_acid, which can tell us that either pH is responsible in the increase of                          citric_acid and fixed_acidity or both of these variables are responsible in the                          decrease in pH in high wine quality.  
                  
## Citation

#### [Domenic V Cicchetti1, Arnold F Cicchetti2. Wine rating scales: Assessing their utility for producers, consumers, and oenologic researchers.Yale Home Office, North Branford, CT, USA; 2Director of National Accounts, Wine Warehouse, San Anselmo, CA, USA](https://pdfs.semanticscholar.org/9ffc/11cd9bbe3f715e0f536fe3b789a60b112397.pdf)
#### [Adding percentage sign in ggplot ](https://stackoverflow.com/questions/40249943/adding-percentage-labels-to-a-bar-chart-in-ggplot2)

